Mire microclimate: groundwater buffers temperature in waterlogged versus dry soils

### 2020-10-26 README.Rmd
zenodo doi
updated readme

@@

# Mire microclimate: groundwater buffers temperature in waterlogged versus dry soils

This repository stores all the information related to the publication *Mire microclimate: groundwater buffers temperature in waterlogged versus dry soils*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the [manuscript](https://github.com/efernandezpascual/mires/blob/master/doc/manuscript.md).

@@

Please cite any use of this repository as `Fernández-Pascual, E. & Correia-Álvarez., E. (2021). Mire microclimate: groundwater buffers temperature in waterlogged versus dry soils. International Journal of Climatology, in press.`

>>>>

[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.4081418.svg)](https://doi.org/10.5281/zenodo.4081418)

@@

# Mire microclimate: Groundwater buffers temperature in waterlogged versus dry soils

This repository stores all the information related to the publication [Mire microclimate: Groundwater buffers temperature in waterlogged versus dry soils](https://doi.org/10.1002/joc.6893), including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.

@@

Please cite any use of this repository as:

> Fernández-Pascual E & Correia-Álvarez E (2021) [Mire microclimate: Groundwater buffers temperature in waterlogged versus dry soils.](https://doi.org/10.1002/joc.6893) *International Journal of Climatology*.

### 2020-10-26 doc/manuscript.Rmd
updated readme

```{r fig1, echo = FALSE, message = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "**Figure 1** **(a)** *Location of the study sites in northern Spain.* **(b)** *Representation of the sampling design at El Riotuertu site. The position of the dry and waterlogged temperature loggers is indicated, the blue area represents the extent of the waterlogged mire habitat.*", fig.width=7, fig.height=3.5}
knitr::include_graphics("../results/Fig1.png")
```

@@

```{r fig2, echo = FALSE, message = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "**Figure 2** *Hourly soil temperature records at the mire sites. The blue series was recorded within the mire, in a waterlogged area. The red series was recorded in a neighbouring dry area. Dataloggers were buried at 5 cm depth. The amplitude of diurnal temperature range in dry sites was reduced with increasing elevation, where the thermal buffer effect was also less apparent.*", fig.width=7, fig.height=10}
knitr::include_graphics("../results/Fig2.png")
```

@@

```{r fig3, echo = FALSE, message = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "**Figure 3** *Average bioclimatic variables in the dry and waterlogged points. The bars represent the mean value, and the brackets the standard error of 8 records.*", fig.width=7, fig.height=3}
knitr::include_graphics("../results/Fig3.png")
```

@@

```{r fig4, echo = FALSE, message = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "**Figure 4** *Scatter plots of the bioclimatic variables predicted by CHELSA air temperatures vs. soil temperatures measured in situ, in dry and waterlogged points.*", fig.width=7, fig.height=7}
knitr::include_graphics("../results/Fig4.png")
```

@@

```{r fig5, echo = FALSE, message = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "**Figure 5** *Complementary temporal analysis of the groundwater buffer effect during peak hot events. Variation of soil temperatures at paired sites (dry vs. waterlogged points) during exceptionally warm days of the warm season. Exceptionally warm days were considered those with a daily maximum (at the dry point) belonging to the 90th percentile of daily maximums for the site. The 1:1 line (corresponding to no buffer effect) is shown, points falling to the right of this line show a buffer effect (i.e. soil temperatures at the waterlogged point are cooler than the neighbouring dry site).*", fig.width=3.5, fig.height=3.5}
knitr::include_graphics("../results/Fig5.png")
```

>>>>

```{r fig1, echo = FALSE, message = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "**Figure 1** **(a)** *Location of the study sites in northern Spain.* **(b)** *Representation of the sampling design at El Riotuertu site. The position of the dry and waterlogged temperature loggers is indicated, the blue area represents the extent of the waterlogged mire habitat.*"}
knitr::include_graphics("../results/Fig1.png")
```

@@

```{r fig2, echo = FALSE, message = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "**Figure 2** *Hourly soil temperature records at the mire sites. The blue series was recorded within the mire, in a waterlogged area. The red series was recorded in a neighbouring dry area. Dataloggers were buried at 5 cm depth. The amplitude of diurnal temperature range in dry sites was reduced with increasing elevation, where the thermal buffer effect was also less apparent.*"}
knitr::include_graphics("../results/Fig2.png")
```

@@

```{r fig3, echo = FALSE, message = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "**Figure 3** *Average bioclimatic variables in the dry and waterlogged points. The bars represent the mean value, and the brackets the standard error of 8 records.*"}
knitr::include_graphics("../results/Fig3.png")
```

@@

```{r fig4, echo = FALSE, message = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "**Figure 4** *Scatter plots of the bioclimatic variables predicted by CHELSA air temperatures vs. soil temperatures measured in situ, in dry and waterlogged points.*"}
knitr::include_graphics("../results/Fig4.png")
```

@@

```{r fig5, echo = FALSE, message = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "**Figure 5** *Complementary temporal analysis of the groundwater buffer effect during peak hot events. Variation of soil temperatures at paired sites (dry vs. waterlogged points) during exceptionally warm days of the warm season. Exceptionally warm days were considered those with a daily maximum (at the dry point) belonging to the 90th percentile of daily maximums for the site. The 1:1 line (corresponding to no buffer effect) is shown, points falling to the right of this line show a buffer effect (i.e. soil temperatures at the waterlogged point are cooler than the neighbouring dry site).*"}
knitr::include_graphics("../results/Fig5.png")
```
